Manuscript

### 2018-12-19 manuscript.Rmd
set up papaja manuscript.Rmd for reproducibility

---
title: "Manuscript"
author: "Sally Claridge"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document:
    theme: cerulean
    number_sections: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
    code_folding: hide
    pandoc_args: [
      "+RTS", "-K2048m",
      "-RTS"
    ]
---

Analysis of the Broad's Avana CRISPR (Broad Institute Cancer Dependency Map 2018, Meyers et al. 2017) and the Broad and Dana-Farber Cancer Institite's Achilles shRNA (MacFarland et al. 2018, Data Science 2018).

The Avana screen produced results using [CERES](https://depmap.org/ceres/) (Meyers et al. 2017) ([GitHub](https://github.com/cancerdatasci/ceres)), which generates gene dependency scores from sgRNA depletion scores from gene essentiality screens and eliminates bias arising from the effect of copy number variation on Cas9 DNA cleavage. The lower the CERES score, the higher the likelihood that the gene is essential in the associated cell line. Scores are scaled per cell line such that a score of 0 is the median effect of nonessential genes and -1 is the median effect of common core essential genes.

All annotation files (copy number, mutation status, and gene expression) (Consortium and Consortium 2015, Barretina et al. 2012) were downloaded from the [DepMap Data Portal](https://depmap.org/portal/download/).

# SET UP

--------------------------------------------------------------------------------

```{r setup, include=FALSE}
set.seed(1234)
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
# num_cores <- parallel::detectCores()
options(mc.cores = parallel::detectCores(), expressions = 5e5)
```

Edited `.Renviron` to avoid "Error: memory exhausted (limit reached?)" when melting the GE dataset, as per the suggestion from [StackOverflow](https://stackoverflow.com/questions/51295402/r-on-macos-error-vector-memory-exhausted-limit-reached)
```
cd ~
touch .Renviron
open .Renviron
# Added following line to .Renviron:
R_MAX_VSIZE=100Gb
```

```{r}
library(bsselectR)
library(cowplot)
theme_set(theme_light())
library(data.table)
library(ggpubr)
library(ggrepel)
library(ggsignif)
library(gplots)
library(gridExtra)
library(htmlwidgets)
library(kableExtra)
library(plotly)
library(reshape2)
library(scales)
library(tidyverse)
library(VennDiagram)
```

# FUNCTIONS {.tabset .tabset-fade .tabset-pills}

--------------------------------------------------------------------------------

## Significance thresholding

```{r}
adj_signif <- function(df) {
  # Takes df from compare_means and creates a signif code column for adjusted p-vals
  df$p.signif <- ifelse(df$p.signif == "ns", NA, df$p.signif)
  df$p.signif.adj <- ifelse(df$p.adj <= 0.0001, "****",
                     ifelse(df$p.adj <= 0.001, "***",
                     ifelse(df$p.adj <= 0.01, "**",
                     ifelse(df$p.adj <= 0.05, "*", NA))))
  df$p.short <- formatC(df$p, format = "g", digits =  2)
  df$p.adj.short <- formatC(df$p.adj, format = "g", digits =  2)

  return(df)
}
```

## Wilcoxon tests by drug

```{r}
WilcoxonByDrug <- function(drug, dataset, data_name, grouping_var) {
  if(grouping_var == "lineage") {
    grouper <- c("Hugo_Symbol", "group_general_lineage_name")
  }
  else if(grouping_var == "gene") {
    grouper <- "Hugo_Symbol"
  }
  use_data <- filter(dataset, Drug == drug)
  sig <- compare_means(AUC_zscore ~ Mutation_Status_Nonsilent, group.by = grouper, data = use_data, method = "wilcox.test", p.adjust.method = "BH")
  sig <- adj_signif(sig)
  sig <- sig[order(sig$p),]
  sig$Drug <- drug
  sig$Dataset <- toupper(data_name)
  return(sig)
}

# WilcoxonByDrugAllGenes <- function(melt_file, grid_file, data_name) {
#   druglist <- as.character(unique(melt_file$Drug))
#   sig_df <- NULL
#   for(i in 1:length(druglist)) {
#     print(paste0("Current drug in process: ", druglist[i], " (drug ", i, " of ", length(druglist), ")."))
#     flush.console()
#     melted <- filter(melt_file, Drug == druglist[i])
#     use_data <- merge(grid_file, melted, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, ccl_converter, by = "accession_id", all.x = TRUE)
#     use_data <- merge(use_data, maf_df, by = c("Hugo_Symbol", "CCLE_Name", "DepMap_ID"), all.x = TRUE)
#     use_data$Mutation_Status_Nonsilent <- ifelse(is.na(use_data$Mutation_Status_Nonsilent), "Wildtype", use_data$Mutation_Status_Nonsilent)
#     use_data$Hugo_Symbol <- factor(use_data$Hugo_Symbol)
#     use_data$AUC <- as.numeric(use_data$AUC)
#     use_data <- filter(use_data, !is.na(use_data$AUC))
#     use_data$Drug_Gene <- paste0(use_data$Drug, "_", use_data$Hugo_Symbol)
#     use_data_ptmuts <- filter(use_data, Reference_Allele_Length == 1 | Reference_Allele_Length == 0 | is.na(Reference_Allele_Length))
#     use_data_ptmuts$Keep_PtMut_Tests <- ifelse(use_data_ptmuts$Reference_Allele_Length == 1 | is.na(use_data_ptmuts$Reference_Allele_Length), TRUE, ifelse(use_data_ptmuts$Tumor_Seq_Allele1_Length == 1, TRUE, FALSE))
#     use_data_ptmuts <- filter(use_data_ptmuts, Keep_PtMut_Tests == TRUE)
#     
#     sig <- compare_means(AUC ~ Mutation_Status_Nonsilent, group.by = "Hugo_Symbol", data = use_data_ptmuts, method = "wilcox.test", p.adjust.method = "BH")
#     sig <- adj_signif(sig)
#     sig <- sig[order(sig$p),]
#     sig <- filter(sig, p <= 0.05)
#     sig$Drug <- druglist[i]
#     sig$Dataset <- toupper(data_name)
#     sig_df <- rbind(sig_df, sig)
#   }
#   return(sig_df)
# }
```

## Pearson correlation by lineage
```{r}
PearsonByLineage <- function(lin, dataset, variable) {
  use_data <- filter(dataset, group_general_lineage_name == lin)
  if(variable == "GE") {
    use_data <- filter(use_data, !is.na(RPKM))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_GE = cor.test(y = AUC_zscore, x = RPKM, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_GE_Abs <- abs(pearson_res$Pearson_Corr_GE)
      pearson_res$Pearson_GE_Sign <- ifelse(sign(pearson_res$Pearson_Corr_GE) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_GE) == -1, "-", "0"))
      return(pearson_res)
    }
  }
  else if(variable == "CN") {
    use_data <- filter(use_data, !is.na(CN))
    use_data_summ <- use_data %>% count(Drug_Gene)
    use_data_summ <- filter(use_data_summ, n >= 3) # cor.test requires at least 3 non-NA data points
    use_data_filt <- filter(use_data, Drug_Gene %in% as.character(unique(use_data_summ$Drug_Gene)))
    if(nrow(use_data_filt) > 0) {
      pearson_res <- use_data_filt %>% group_by(Drug_Gene) %>% summarize(Pearson_Corr_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$estimate, Pearson_Pval_CN = cor.test(y = AUC_zscore, x = CN, method = "pearson", use = "complete.obs")$p.value)
      pearson_res <- merge(g2p_druggene_tally, pearson_res, all.y = TRUE)
      pearson_res$group_general_lineage_name <- lin
      pearson_res$Pearson_CN_Abs <- abs(pearson_res$Pearson_Corr_CN)
      pearson_res$Pearson_CN_Sign <- ifelse(sign(pearson_res$Pearson_Corr_CN) == 1, "+",
                                              ifelse(sign(pearson_res$Pearson_Corr_CN) == -1, "-", "0"))
      return(pearson_res)
    }
  }
}
```

## Save widget to new directory
```{r}
saveWidgetFix <- function (widget , file, ...) {
  ## A wrapper to saveWidget which compensates for arguable BUG in
  ## saveWidget which requires `file` to be in current working
  ## directory.
  wd <- getwd()
  on.exit(setwd(wd))
  outDir <- dirname(file)
  file <- basename(file)
  setwd(outDir);
  saveWidget(widget, file = file,...)
}
```

## Make lineage mut plots

```{r}
MutPlotsByLineage <- function(lineage_name) {

  use_data <- filter(dgl_mut_grid_summ, group_general_lineage_name == lineage_name)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  ticker <- 0

  use_data_ccle <- filter(use_data, Dataset == "CCLE")
  use_data_ccle_summ <- use_data_ccle %>% count(InG2P)
  ccle_lin_p_ttest <- data.frame(statistic = NA,
                                 parameter =  NA,
                                 p.value = NA)
  if(nrow(use_data_ccle_summ) <= 1) {
    ccle_lin_p_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn < 2) { ccle_lin_p_ttest_res <-  NA }
  else if(use_data_ccle_summ[use_data_ccle_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_lin_p_ttest <- t.test(use_data_ccle$p ~ use_data_ccle$InG2P)
    ccle_lin_p_ttest_res <- paste0("t(", trunc(ccle_lin_p_ttest$parameter), ") = ", round(ccle_lin_p_ttest$statistic, 2), ", p = ", round(ccle_lin_p_ttest$p.value, 3))
  }

  use_data_ctrp <- filter(use_data, Dataset == "CTRP")
  use_data_ctrp_summ <- use_data_ctrp %>% count(InG2P)
  ctrp_lin_p_ttest <- data.frame(statistic = NA,
                                 parameter =  NA,
                                 p.value = NA)
  if(nrow(use_data_ctrp_summ) <= 1) {
    ctrp_lin_p_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn < 2) { ctrp_lin_p_ttest_res <-  NA }
  else if(use_data_ctrp_summ[use_data_ctrp_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_lin_p_ttest <- t.test(use_data_ctrp$p ~ use_data_ctrp$InG2P)
    ctrp_lin_p_ttest_res <- paste0("t(", trunc(ctrp_lin_p_ttest$parameter), ") = ", round(ctrp_lin_p_ttest$statistic, 2), ", p = ", round(ctrp_lin_p_ttest$p.value, 3))
  }

  use_data_gdsc <- filter(use_data, Dataset == "GDSC")
  use_data_gdsc_summ <- use_data_gdsc %>% count(InG2P)
  gdsc_lin_p_ttest <- data.frame(statistic = NA,
                                 parameter =  NA,
                                 p.value = NA)
  if(nrow(use_data_gdsc_summ) <= 1) {
    gdsc_lin_p_ttest_res <-  NA
    ticker <- ticker + 1
    }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn < 2) { gdsc_lin_p_ttest_res <-  NA }
  else if(use_data_gdsc_summ[use_data_gdsc_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_lin_p_ttest <- t.test(use_data_gdsc$p ~ use_data_gdsc$InG2P)
    gdsc_lin_p_ttest_res <- paste0("t(", trunc(gdsc_lin_p_ttest$parameter), ") = ", round(gdsc_lin_p_ttest$statistic, 2), ", p = ", round(gdsc_lin_p_ttest$p.value, 3))
  }

  y_value <- ifelse(ticker == 3, 10, max(use_data$n) + max(use_data$n) * 0.1)

  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                          Test_Results_p = c(ccle_lin_p_ttest_res, ctrp_lin_p_ttest_res, gdsc_lin_p_ttest_res),
                          t_statistic = c(round(ccle_lin_p_ttest$statistic, 3), round(ctrp_lin_p_ttest$statistic, 3), round(gdsc_lin_p_ttest$statistic, 3)),
                          df = c(trunc(ccle_lin_p_ttest$parameter), trunc(ctrp_lin_p_ttest$parameter), trunc(gdsc_lin_p_ttest$parameter)),
                          p_value = c(round(ccle_lin_p_ttest$p.value, 3), round(ctrp_lin_p_ttest$p.value, 3), round(gdsc_lin_p_ttest$p.value, 3)),
                          group_general_lineage_name = lineage_name)
  plot_point <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(xlim = c(0, 1), ylim = c(0, y_value)) +
    scale_x_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    scale_y_continuous(breaks = pretty_breaks()) +
    geom_point(mapping = aes(label = Drug_Gene, color = InG2P, x = p, y = n), alpha = 0.5) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    geom_text_repel(data = subset(use_data, InG2P == "Yes"), mapping = aes(x = p, y = n, label = Drug_Gene), size = 3, segment.size = 0.2, segment.color = "grey50") +
    geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value, label = Test_Results_p), nudge_y = 0.1) +
    labs(x = "Drug-gene association p-value between mutation groups", y = "Number of level A G2P associations") +
    theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
  plot_box <- ggplot(data = use_data) +
    facet_wrap(~ Dataset, drop = FALSE) +
    coord_cartesian(ylim = c(0, 1)) +
    scale_y_continuous(breaks = seq(0, 1, by = 0.1), labels = seq(0, 1, by = 0.1)) +
    geom_violin(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.5) +
    geom_boxplot(mapping = aes(color = InG2P, y = p, x = InG2P), alpha = 0.75, width = 0.25) +
    scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
    coord_flip() +
    labs(x = "In G2P?", title = toupper(lineage_name)) +
    theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))
  plot <- subplot(plot_box, plot_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>% layout(title = toupper(lineage_name), margin = list(t = 75, l = 75), xaxis5 = list(title = "Gene-drug association p-value between mutation groups"), yaxis = list(title = "Level A G2P\nassociation", domain = c(0.8, 1)), yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

  plot_point <- ggplot_gtable(ggplot_build(plot_point))
  plot_box <- ggplot_gtable(ggplot_build(plot_box))
  p_maxWidth <-  grid::unit.pmax(plot_point$widths, plot_box$widths)
  plot_point$widths <- as.list(p_maxWidth)
  plot_box$widths <- as.list(p_maxWidth)

  mut_plot <- cowplot::plot_grid(plot_box,
                     plot_point,
                     nrow = 2,
                     rel_heights = c(1, 4))

  saveWidgetFix(widget = plot, file = paste0("./plots/manuscript/plotlys_mut_by_lineage_grid/", lineage_name, "_mut_plotly_grid.html"))
  ggsave(paste0("./plots/manuscript/pngs_mut_by_lineage_grid/", lineage_name, "_mut_plot_grid.png"), mut_plot, device = "png", dpi = 450, width = 12, height = 6, units =   "in")

  return(summ_text)
}
```

## Make lineage GE/CN plots

```{r}
PearsonPlotsByLineage <- function(lineage) {
  print(paste0("Lineage: ", toupper(lineage)))
  use_data <- filter(dgl_signif_pearson_g2p_grid, group_general_lineage_name == lineage)
  use_data$Dataset <- factor(use_data$Dataset, levels = c("CCLE", "CTRP", "GDSC"))
  use_data$Pearson_GE_Label <- paste0(round(use_data$Pearson_Pval_GE, 3), "\nDrug_Gene: ", use_data$Drug_Gene)
  use_data$Pearson_CN_Label <- paste0(round(use_data$Pearson_Pval_CN, 3), "\nDrug_Gene: ", use_data$Drug_Gene)
  ticker_ge <- 0
  ticker_cn <- 0
  ticker <- 0

  use_data_ccle <- filter(use_data, Dataset == "CCLE")
  use_data_ccle_ge_summ <- use_data_ccle[!is.na(use_data_ccle$Pearson_GE_Abs),] %>% count(InG2P)
  use_data_ccle_cn_summ <- use_data_ccle[!is.na(use_data_ccle$Pearson_CN_Abs),] %>% count(InG2P)
  ccle_ge_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  ccle_cn_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  if(nrow(use_data_ccle_ge_summ) <= 1) {
    ccle_ge_res <-  NA
    ticker_ge <- ticker_ge + 1
  }
  else if(use_data_ccle_ge_summ[use_data_ccle_ge_summ$InG2P == "Yes",]$nn < 2) {
    ccle_ge_res <-  NA
  }
  else if(use_data_ccle_ge_summ[use_data_ccle_ge_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_ge_ttest <- t.test(use_data_ccle$Pearson_GE_Abs ~ use_data_ccle$InG2P)
    ccle_ge_res <- paste0("t(", trunc(ccle_ge_ttest$parameter), ") = ", round(ccle_ge_ttest$statistic, 2), ", p = ", round(ccle_ge_ttest$p.value, 3))
  }
  if(nrow(use_data_ccle_cn_summ) <= 1) {
    ccle_cn_res <-  NA
    ticker_cn <- ticker_cn + 1
  }
  else if(use_data_ccle_cn_summ[use_data_ccle_cn_summ$InG2P == "Yes",]$nn < 2) {
    ccle_cn_res <-  NA
  }
  else if(use_data_ccle_cn_summ[use_data_ccle_cn_summ$InG2P == "Yes",]$nn >= 2) {
    ccle_cn_ttest <- t.test(use_data_ccle$Pearson_CN_Abs ~ use_data_ccle$InG2P)
    ccle_cn_res <- paste0("t(", trunc(ccle_cn_ttest$parameter), ") = ", round(ccle_cn_ttest$statistic, 2), ", p = ", round(ccle_cn_ttest$p.value, 3))
  }
  print("CCLE testing: DONE")

  use_data_ctrp <- filter(use_data, Dataset == "CTRP")
  use_data_ctrp_ge_summ <- use_data_ctrp[!is.na(use_data_ctrp$Pearson_GE_Abs),] %>% count(InG2P)
  use_data_ctrp_cn_summ <- use_data_ctrp[!is.na(use_data_ctrp$Pearson_CN_Abs),] %>% count(InG2P)
  ctrp_ge_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  ctrp_cn_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  if(nrow(use_data_ctrp_ge_summ) <= 1) {
    ctrp_ge_res <-  NA
    ticker_ge <- ticker_ge + 1
  }
  else if(use_data_ctrp_ge_summ[use_data_ctrp_ge_summ$InG2P == "Yes",]$nn < 2) {
    ctrp_ge_res <-  NA
  }
  else if(use_data_ctrp_ge_summ[use_data_ctrp_ge_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_ge_ttest <- t.test(use_data_ctrp$Pearson_GE_Abs ~ use_data_ctrp$InG2P)
    ctrp_ge_res <- paste0("t(", trunc(ctrp_ge_ttest$parameter), ") = ", round(ctrp_ge_ttest$statistic, 2), ", p = ", round(ctrp_ge_ttest$p.value, 3))
  }
  if(nrow(use_data_ctrp_cn_summ) <= 1) {
    ctrp_cn_res <-  NA
    ticker_cn <- ticker_cn + 1
  }
  else if(use_data_ctrp_cn_summ[use_data_ctrp_cn_summ$InG2P == "Yes",]$nn < 2) {
    ctrp_cn_res <-  NA
  }
  else if(use_data_ctrp_cn_summ[use_data_ctrp_cn_summ$InG2P == "Yes",]$nn >= 2) {
    ctrp_cn_ttest <- t.test(use_data_ctrp$Pearson_CN_Abs ~ use_data_ctrp$InG2P)
    ctrp_cn_res <- paste0("t(", trunc(ctrp_cn_ttest$parameter), ") = ", round(ctrp_cn_ttest$statistic, 2), ", p = ", round(ctrp_cn_ttest$p.value, 3))
  }
  print("CTRP testing: DONE")

  use_data_gdsc <- filter(use_data, Dataset == "GDSC")
  use_data_gdsc_ge_summ <- use_data_gdsc[!is.na(use_data_gdsc$Pearson_GE_Abs),] %>% count(InG2P)
  use_data_gdsc_cn_summ <- use_data_gdsc[!is.na(use_data_gdsc$Pearson_CN_Abs),] %>% count(InG2P)
  gdsc_ge_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  gdsc_cn_ttest <- data.frame(statistic = NA, parameter = NA, p.value = NA)
  if(nrow(use_data_gdsc_ge_summ) <= 1) {
    gdsc_ge_res <-  NA
    ticker_ge <- ticker_ge + 1
  }
  else if(use_data_gdsc_ge_summ[use_data_gdsc_ge_summ$InG2P == "Yes",]$nn < 2) {
    gdsc_ge_res <-  NA
  }
  else if(use_data_gdsc_ge_summ[use_data_gdsc_ge_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_ge_ttest <- t.test(use_data_gdsc$Pearson_GE_Abs ~ use_data_gdsc$InG2P)
    gdsc_ge_res <- paste0("t(", trunc(gdsc_ge_ttest$parameter), ") = ", round(gdsc_ge_ttest$statistic, 2), ", p = ", round(gdsc_ge_ttest$p.value, 3))
  }
  if(nrow(use_data_gdsc_cn_summ) <= 1) {
    gdsc_cn_res <-  NA
    ticker_cn <- ticker_cn + 1
  }
  else if(use_data_gdsc_cn_summ[use_data_gdsc_cn_summ$InG2P == "Yes",]$nn < 2) {
    gdsc_cn_res <-  NA
  }
  else if(use_data_gdsc_cn_summ[use_data_gdsc_cn_summ$InG2P == "Yes",]$nn >= 2) {
    gdsc_cn_ttest <- t.test(use_data_gdsc$Pearson_CN_Abs ~ use_data_gdsc$InG2P)
    gdsc_cn_res <- paste0("t(", trunc(gdsc_cn_ttest$parameter), ") = ", round(gdsc_cn_ttest$statistic, 2), ", p = ", round(gdsc_cn_ttest$p.value, 3))
  }
  print("GDSC testing: DONE")

  y_value_ge <- ifelse(ticker_ge == 3, 10, max(use_data$n) + max(use_data$n) * 0.1)
  y_value_cn <- ifelse(ticker_cn == 3, 10, max(use_data$n) + max(use_data$n) * 0.1)

  summ_text <- data.frame(Dataset = c("CCLE", "CTRP", "GDSC"),
                          Test_Results_GE = c(ccle_ge_res, ctrp_ge_res, gdsc_ge_res),
                          t_statistic_GE = c(round(ccle_ge_ttest$statistic, 3), round(ctrp_ge_ttest$statistic, 3), round(gdsc_ge_ttest$statistic, 3)),
                          df_GE = c(trunc(ccle_ge_ttest$parameter), trunc(ctrp_ge_ttest$parameter), trunc(gdsc_ge_ttest$parameter)),
                          p_value_GE = c(round(ccle_ge_ttest$p.value, 3), round(ctrp_ge_ttest$p.value, 3), round(gdsc_ge_ttest$p.value, 3)),
                          Test_Results_CN = c(ccle_cn_res, ctrp_cn_res, gdsc_cn_res),
                          t_statistic_CN = c(round(ccle_cn_ttest$statistic, 3), round(ctrp_cn_ttest$statistic, 3), round(gdsc_cn_ttest$statistic, 3)),
                          df_CN = c(trunc(ccle_cn_ttest$parameter), trunc(ctrp_cn_ttest$parameter), trunc(gdsc_cn_ttest$parameter)),
                          p_value_CN = c(round(ccle_cn_ttest$p.value, 3), round(ctrp_cn_ttest$p.value, 3), round(gdsc_cn_ttest$p.value, 3)),
                          group_general_lineage_name = lineage)

  ## Gene expression
  if(sum(!is.na(use_data$Pearson_GE_Abs)) > 0) {
    plot_ge_point <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      scale_y_continuous(breaks = pretty_breaks()) +
      geom_point(mapping = aes(label = Pearson_GE_Label, color = InG2P, fill = InG2P, shape = Pearson_GE_Sign, x = Pearson_GE_Abs, y = n), alpha = 0.5) +
      scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      geom_text_repel(data = subset(use_data, InG2P == "Yes" & n > 10), aes(label = Drug_Gene, x = Pearson_GE_Abs, y = n), size = 3, segment.size = 0.2, segment.color = "grey50") +
      geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value_ge, label = Test_Results_GE), nudge_y = 1) +
      labs(y = "Number of level A G2P associations", x = "Pearson correlation coefficient between drug-gene AUC and gene expression") +
      theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
    plot_ge_box <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      geom_violin(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.5) +
      geom_boxplot(mapping = aes(color = InG2P, y = Pearson_GE_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      coord_flip() +
      labs(x = "In G2P?", title = toupper(lineage)) +
      theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

    plot_ge <- subplot(plot_ge_box, plot_ge_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>%
               layout(title = lineage, margin = list(t = 75, l = 75),
                      xaxis5 = list(title = "Pearson correlation coefficient between drug-gene AUC and gene expression"),
                      yaxis = list(title = "In G2P?", domain = c(0.8, 1)),
                      yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

    png_ge_point <- ggplot_gtable(ggplot_build(plot_ge_point))
    png_ge_box <- ggplot_gtable(ggplot_build(plot_ge_box))
    p_maxWidth <-  grid::unit.pmax(png_ge_point$widths, png_ge_box$widths)
    png_ge_point$widths <- as.list(p_maxWidth)
    png_ge_box$widths <- as.list(p_maxWidth)

    png_ge <- cowplot::plot_grid(png_ge_box, png_ge_point, nrow = 2, rel_heights = c(1, 4))
    png_ge

    saveWidgetFix(widget = plot_ge, file = paste0("./plots/manuscript/plotlys_ge_by_lineage_grid/", lineage, "_ge_plotly_grid.html"))
    ggsave(paste0("./plots/manuscript/pngs_ge_by_lineage_grid/", lineage, "_ge_plot_grid.png"), png_ge, device = "png", dpi = 450, width = 12, height = 6, units = "in")
    print("Gene expression plots: DONE")
  }
  else if(sum(!is.na(use_data$Pearson_GE_Abs)) == 0) { print("Gene expression plots: None") }

  ## Copy number
  if(sum(!is.na(use_data$Pearson_CN_Abs)) > 0) {
    plot_cn_point <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      scale_y_continuous(breaks = pretty_breaks()) +
      geom_point(mapping = aes(label = Pearson_CN_Label, color = InG2P, fill = InG2P, shape = Pearson_CN_Sign, x = Pearson_CN_Abs, y = n), alpha = 0.5, position = position_jitter(width = 0, height = 0.15)) +
      scale_shape_manual(values = c(25, 24), labels = c("Negative", "Positive")) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      scale_fill_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      geom_text_repel(data = subset(use_data, InG2P == "Yes"  & n > 10), aes(label = Drug_Gene, x = Pearson_CN_Abs, y = n), size = 3, segment.size = 0.2, segment.color = "grey50") +
      geom_text(data = summ_text, mapping = aes(x = 0.5, y = y_value_cn, label = Test_Results_CN), nudge_y = 1) +
      labs(y = "Number of level A G2P associations", x = "Pearson correlation coefficient between drug-gene AUC and copy number") +
      theme(legend.position = "none", plot.margin = unit(c(0.05, 0.5, 0.5, 0.5), "cm"), strip.background = element_blank(), strip.text.x = element_blank())
    plot_cn_box <- ggplot(data = use_data) +
      facet_wrap(~ Dataset, drop = FALSE) +
      geom_violin(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.5) +
      geom_boxplot(mapping = aes(color = InG2P, y = Pearson_CN_Abs, x = InG2P), alpha = 0.5, width = 0.25) +
      scale_color_manual(values = c("Yes" = "darkorchid4", "No" = "lightpink")) +
      coord_flip() +
      labs(x = "In G2P?", title = toupper(lineage)) +
      theme(legend.position = "none", axis.title.x = element_blank(), axis.text.x = element_blank(), axis.ticks.x = element_blank(), plot.margin = unit(c(0.5, 0.5, 0.05, 0.5), "cm"))

    plot_cn <- subplot(plot_cn_box, plot_cn_point, nrows = 2, titleX = TRUE, titleY = TRUE) %>%
               layout(title = lineage, margin = list(t = 75, l = 75),
                      xaxis5 = list(title = "Pearson correlation coefficient between drug-gene AUC and copy number"),
                      yaxis = list(title = "In G2P?", domain = c(0.8, 1)),
                      yaxis2 = list(title = "Number of level A G2P associations", domain = c(0, 0.79)))

    png_cn_point <- ggplot_gtable(ggplot_build(plot_cn_point))
    png_cn_box <- ggplot_gtable(ggplot_build(plot_cn_box))
    p_maxWidth <-  grid::unit.pmax(png_cn_point$widths, png_cn_box$widths)
    png_cn_point$widths <- as.list(p_maxWidth)
    png_cn_box$widths <- as.list(p_maxWidth)

>>>>

---
title             : "Manuscript"
shorttitle        : "Subtitle"

author: 
  - name          : "Sally E. Claridge"
    affiliation   : "1"
  - name          : "Daniel M. Charytonowicz"
    affiliation   : "1"
  - name          : "Adam A. Margolin"
    affiliation   : "1, 2"
    corresponding : yes    # Define only one corresponding author
    address       : "Icahn School of Medicine at Mount Sinai, Deparment of Genetics and Genomic Sciences, One Gustave L. Levy Place, Box 1498, New York, NY 10029"
    email         : "adam.margolin@mssm.edu"

affiliation:
  - id            : "1"
    institution   : "Department of Genetics and Genomic Sciences, Icahn Institute for Data Science and Genomic Technology, Icahn School of Medicine at Mount Sinai, New York, NY"
  - id            : "2"
    institution   : "Cancer Target Discovery and Development Network (Oregon Health and Science University, Portland, OR), National Cancer Institute, National Institutes of Health"

authornote: |
  Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

  Enter author note here.

abstract: |
  One or two sentences providing a **basic introduction** to the field,  comprehensible to a scientist in any discipline.

  Two to three sentences of **more detailed background**, comprehensible  to scientists in related disciplines.

  One sentence clearly stating the **general problem** being addressed by  this particular study.

  One sentence summarizing the main result (with the words "**here we show**" or their equivalent).

  Two or three sentences explaining what the **main result** reveals in direct comparison to what was thought to be the case previously, or how the  main result adds to previous knowledge.

  One or two sentences to put the results into a more **general context**.

  Two or three sentences to provide a **broader perspective**, readily comprehensible to a scientist in any discipline.

  <!-- https://tinyurl.com/ybremelq -->

keywords          : "keywords"
wordcount         : "X"

bibliography      : ["manuscript.bib"]

floatsintext      : no
figurelist        : no
tablelist         : no
footnotelist      : no
linenumbers       : yes
mask              : no
draft             : no

documentclass     : "apa6"
classoption       : "man"
biblio-style      : "apa"
output            :
  papaja::apa6_pdf:
    citation_package: biblatex
---

```{r setup, include = FALSE}
library("papaja")
# install.packages("tinytex")
# tinytex::install_tinytex()
# install.packages("citr")
```

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

[@aksoy2017]

# Introduction

# Results

# Discussion

\newpage

# Methods

We used `r cite_r("manuscript_r_references.bib")` for all our analyses.

\newpage

# References

```{r create_r-references}
r_refs(file = "manuscript_r_references.bib")
```

\begingroup
\setlength{\parindent}{-0.5in}
\setlength{\leftskip}{0.5in}

Finally, the APA guidelines require a note at the start of the reference section that explains what an asterisk means. This note can be added at the end of the document as follows.

<div id = "refs"></div>
\endgroup
